---
title: "Week7HypothesisTesting"
author: "Anusha Gadgil"
date: "2023-10-07"
output: html_document
---

#Loading the packages
```{r}
library(tidyverse)
library(dplyr)
library(reshape2)
library(gridExtra)
library(ggplot2)

# Install the pwr package if you haven't already
install.packages("pwr")

# Load the pwr package
library(pwr)
```

## Loading the Dataset

```{r}
WineDataset<-read.csv("C:/Users/Anusha Gadgil/Desktop/wines_SPA (1).csv")
head(WineDataset)
sample_size<-nrow(WineDataset)
print(sample_size)
```

## FIRST NULL HYPOTHESIS
##1. There is no relationship between the Price and rating of Wines
```{r}

#I picked an alpha value of 0.05 because it is the standard recommended value 
alpha<-0.05
#I randomly selected a power level and then used that selected level to run a power test and check the required sample size for the given power level
power <- 0.80

min_effect_size <- 0.2

# Check if the p-value is less than alpha

```

#checking if I have enough data to perform neyman pearsons test

```{r}
effect_size <- 0.04 
alpha <- 0.05      
power <- 0.80       


required_sample_size <- pwr.t.test(
  d = effect_size,
  sig.level = alpha,
  power = power,
  type = "two.sample"  
)$n


if (required_sample_size >= sample_size) {
  cat("You have enough data to perform the Neyman-Pearson test.\n")
  # Perform your hypothesis test here
} else {
  cat("You do not have enough data to perform the Neyman-Pearson test.\n")
  cat("Consider increasing your sample size to achieve the desired power.\n")
}

# Print the required sample size
print(required_sample_size)

```


#Performing Neyman Pearson's test

```{r}
library(readr)
# Extract the two columns
column1 <- WineDataset$rating
column2 <- WineDataset$price



# Calculate sample statistics
mean1 <- mean(column1)
mean2 <- mean(column2)
n1 <- length(column1)
n2 <- length(column2)

# Set alpha(significance level)
alpha <- 0.05  

# Calculate the test statistic (assuming equal variances)
pooled_variance <- ((n1 - 1) * var(column1) + (n2 - 1) * var(column2)) / (n1 + n2 - 2)
t_stat <- (mean1 - mean2) / (sqrt(pooled_variance * (1 / n1 + 1 / n2)))

# Calculate the critical value (assuming a one-tailed test)
critical_value <- qt(1 - alpha, df = n1 + n2 - 2)

# Perform the Neyman-Pearson test
if (t_stat > critical_value) {
  cat("Reject the null hypothesis in favor of the alternative hypothesis.\n")
} else {
  cat("Fail to reject the null hypothesis.\n")
}
```
#Fishers test
```{r}
test_result <- var.test(WineDataset$rating, WineDataset$price)

# Extract the p-value
p_value <- test_result$p.value

# Interpret the p-value
if (p_value < 0.05) {
  cat("The p-value is less than 0.05, so there is evidence to reject the null hypothesis.\n")
  cat("The variances of the two datasets are significantly different.\n")
} else {
  cat("The p-value is greater than or equal to 0.05, so we fail to reject the null hypothesis.\n")
  cat("There is no strong evidence that the variances are different.\n")
}

# Print the test result summary
print(test_result)

```







##SECOND NULL HYPOTHESIS
##2.There is no difference between the average price of old and new wines.



#finding range of the years
```{r}

WineDataset<-read.csv("C:/Users/Anusha Gadgil/Desktop/wines_SPA (1).csv")
head(WineDataset)
WineDataset <- WineDataset[complete.cases(WineDataset[, "year"]), ]

year_range <- range(WineDataset$year)

# Extract the minimum and maximum years
min_year <- year_range[1]
max_year <- year_range[2]

# Print the result
cat("Minimum year:", min_year, "\n")
cat("Maximum year:", max_year, "\n")
print(year_range)
```

#arranging data in order of year
```{r}
wine_data <- WineDataset %>% arrange(year)

# View the resulting dataset
head(wine_data)

```

#Slicing arranged dataset into 2 halfs
```{r}
num_rows <- nrow(wine_data)

# Calculate the midpoint to split the data into two halves
midpoint <- num_rows / 2

# Split the data into two halves
first_half <- wine_data[1:midpoint, ]
second_half <- wine_data[(midpoint + 1):num_rows, ]
print(num_rows)
first_half
```


#Column 1 has the older wine, column 2 has the newer wine
```{r}
Old_price <- first_half$price
New_price<- second_half$price
samplesize<-nrow(first_half)
samplesize
```


#checking if I have enough data in the columns to perform neyman pearsons test

```{r}
effect_size <- 0.04  # Desired effect size ()
alpha <- 0.05       # Significance level (e.g., 0.05 for 5%)
power <- 0.80       # Desired power level (e.g., 0.80 for 80%)


required_sample_size <- pwr.t.test(
  d = effect_size,
  sig.level = alpha,
  power = power,
  type = "two.sample"  # Specify a two-sample t-test
)$n


if (required_sample_size >= samplesize) {
  cat("You have enough data to perform the Neyman-Pearson test.\n")
  # Perform your hypothesis test here
} else {
  cat("You do not have enough data to perform the Neyman-Pearson test.\n")
  cat("Consider increasing your sample size to achieve the desired power.\n")
}

# Print the required sample size
print(required_sample_size)

```
#Performing Neyman Pearson's test
```{r}
library(readr)
# Extract the two columns
Old_price <- first_half$price
New_price<- second_half$price
# Define your null and alternative hypotheses
# H0: μ1 <= μ2 (Mean of column1 is less than or equal to the mean of column2)
# H1: μ1 > μ2 (Mean of column1 is greater than the mean of column2)

# Calculate sample statistics
mean1 <- mean(Old_price)
mean2 <- mean(New_price)
n1 <- length(Old_price)
n2 <- length(New_price)


alpha <- 0.05  
# Calculate the test statistic 
pooled_variance <- ((n1 - 1) * var(column1) + (n2 - 1) * var(column2)) / (n1 + n2 - 2)
t_stat <- (mean1 - mean2) / (sqrt(pooled_variance * (1 / n1 + 1 / n2)))

# Calculating  the critical value 
critical_value <- qt(1 - alpha, df = n1 + n2 - 2)

# Perform the Neyman-Pearson test
if (t_stat > critical_value) {
  cat("Reject the null hypothesis and pick the alternative hypothesis.\n")
} else {
  cat("Fail to reject the null hypothesis.\n")
}
```



#Fishers test
```{r}
test_result <- var.test(first_half$price, second_half$price)

# Extract the p-value
p_value <- test_result$p.value

# Interpret the p-value
if (p_value < 0.05) {
  cat("The p-value is less than 0.05, so reject the null hypothesis.\n")
  cat("The variances of the two datasets are different.\n")
} else {
  cat("The p-value is greater than 0.05, so we fail to reject the null hypothesis.\n")
  cat("There is no strong evidence that the variances are different.\n")
}


print(test_result)

```
#Building a visualisation for Hypothesis 1
```{r}
hist(column1, col = "blue", main = "Histogram of Rating")
hist(column2, col = "red", main = "Histogram of Price")
```
#Building a visualisation for Hypothesis 2
```{r}
boxplot(Old_price, New_price, names = c("Old_price", "New_price"))

```
```{r}
hist(Old_price, col = "blue", main = "Histogram of Old_Price")
hist(New_price, col = "red", main = "Histogram of New_Price")
```